### 2021-03-12 manuscript/manuscript-speed-accuracy.tex
Major revision for manuscript and figures after comments from Mik.

\PaperTitle{A meta-analysis of bioinformatics software benchmarks reveals that publication-bias unduly influences software accuracy} % Article title

@@

%\Abstract{In the below we provide additional results for our investigation of computational biology benchmarks.}
\Abstract{ Computational biology has provided widely used and powerful
  software tools for testing and making inferences about biological
  data. In the face of rapidly increasing volumes of data, heuristic
  methods that trade software speed for accuracy must
  be employed.  We are interested in testing for trade-offs between
  speed and accuracy and for other factors that are indicative of accurate
  software.

We have extracted accuracy and speed ranks from independent benchmarks of
computational biology software of different software tools, and evaluated the
factors that are likely to influence accuracy e.g. speed, author
reputation, journal impact, recency or developer efforts.

We found that software speed, author reputation, journal impact, the
number of citations and age are all unreliable predictors of software
accuracy. This is unfortunate because citations, author and journal
reputation are frequently used reasons for selecting software
tools. Our results do show that accurate bioinformatic software tools
are generally the product of many gradual improvements, often from
multiple developers, based upon matching GitHub records with benchmark results. 
In addition, we find
that the field of bioinformatics has an excess of slow and inaccurate software
tools, this is consistent across many sub-disciplines. Meanwhile, there are few tools that
are middle-of-road in terms of accuracy and speed trade-offs. We
hypothesise that a form of publication-bias unduly influences the
publication and development of bioinformatic software. In other
words, software that is intermediate in terms of both speed and accuracy may be difficult to publish due to author,
editor and reviewer practices. This leaves an unfortunate gap in the
literature as the ideal tools may fall in the gap e.g. high accuracy tools are not always useful if years of CPU time are required, 
while high speed is not useful if the results are also inaccurate.  }

@@

\section*{Background}
Computational biology software is widely used and has produced some of
the most cited publications in the entire scientific corpus
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. This
highly-cited software includes implementations of methods for sequence
alignment and homology inference
\cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
phylogenetic analysis
\cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei},
statistical analysis of survival patterns in biomedicine
\cite{Kaplan1958-ju,Cox1972-nu}, biomolecular structure analysis
\cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj},
visualization and data collection
\cite{Kraulis1991-lt,Berman2000-to}. However, the popularity of a
software tool does not necessarily mean that it is accurate or
computationally efficient, instead usability, ease of installation,
operating system and other factors may play a greater role in a software
tool's popularity. Indeed, there have been several notable incidences where
inaccurate software has caused considerable harm
\cite{leveson1993investigation,cummings2020regulating,Herkert:2020}.
%Fergusson covid predictions?

Progress in the biological sciences is increasingly limited by the
ability to analyse large volumes of data, therefore the
dependence of biologists on software is also increasing
\cite{Marx2013-zi}. There is an increasing reliance on technological
solutions for automating biological data generation
(e.g. next-generation sequencing, mass-spectroscopy, cell-tracking and
species tracking), therefore the biological sciences have become
increasingly dependent upon computational software for processing
large quantities of data \cite{Marx2013-zi}. As a consequence, the
computational efficiency of data processing and analysis software is
of great importance to decrease the energy, climate impact, and time costs of research
\cite{Gombiner2011-md}. Furthermore, even small error rates can have
major impacts on the number of false inferences as datasets become
larger \cite{Storey2003-cv}.

The gold-standard for determining accuracy is for independent
researchers to conduct benchmarks, which can serve a useful role in
reducing the over-optimistic reporting of software accuracy
\cite{Boulesteix2010-te,Jelizarow2010-zf,Weber:2019} and the self-assessment trap
\cite{Norel2011-cq,Buchka:2021}. Benchmark studies typically use a number of
positive and negative control datasets, predictions from different software tools can then be
partitioned into true or false groups and a variety of metrics can be
used to evaluate the performance differences
\cite{Egan1975-nd,Hall2012-kg,Weber:2019}.
%Some benchmarks now use live, or
%frequently updated, results to indicate the latest developments in
%software performance
%\cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}.
The aim of these benchmarks is to independently identify tools that
make acceptable compromises in terms of balancing speed with
discriminating true and false predictions, and are therefore suited
for wide adoption by the community.

For common computational biology tasks, a proliferation of
software-based solutions often exists
\cite{Felsenstein1995-ic,Altschul2013-bv,Henry2014-ut}. While
this is a good problem to have, and points to a
diversity of options from which practical solutions can be selected,
having many possible options creates a dilemma for users. In the absence of
any recent gold-standard benchmarks, how should scientific software be
selected? In the following we presume that ``biological
accuracy'' is the most desirable feature for a software tool. Biological
accuracy is the degree to which predictions or measurements reflect
the biological truths based on expert-derived curated datasets (see Methods for the  mathematical definition used here). In some
fields biological accuracy is very difficult to ascertain, for
example, in phylogenetics it is nearly impossible in most cases to know the
true ancestral relationships between organisms. In situations like this,
researchers may use simulated or high-confidence datasets.

A number of possible predictors of software quality are used by the
community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. Some accessible,
quantifiable and frequently used proxies for identifying high quality
software include: \textbf{1. Recency:} recently published software
tools may have built upon the results of past work, or be an update to
an existing software. Therefore these may be more accurate and
faster. \textbf{2. Wide adoption:} a software tool may be widely used
because it is fast and accurate, or because it is well-supported and
user-friendly. In fact, “large user base”, “word-of-mouth”,
“wide-adoption”, "personal recommendation," and "recommendation from a
close colleague," are frequent responses to surveys of “how do
scientists select software?”
\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
  impact:} high profile journals are run by editors and reviewers who
carefully select and curate the best manuscripts. Therefore, high
impact journals may be more likely to select manuscripts describing
good software \cite{Garfield1955-wf}. \textbf{4. Author/Group
  reputation:} the key to any project is the skills of the people
involved, including maintaining a high collective intelligence
\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
consequence, an argument could be made that well respected and
high-profile authors will produce better software
\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
frequently trade accuracy for speed. For example, heuristic
software such as the popular homology search tool, BLAST, compromises
the mathematical guarantee of optimal solutions for more speed
\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
interpret this fact as slower software is likely to be more
accurate. But speed may also be influenced by the programming language
\cite{Fourment2008-vl}, and the level of hardware optimisation
\cite{Farrar2007-ky,Dematte2010-ph};
however, the implementation generally has more of an impact (e.g. brute-force approaches versus
rapid and sensitive pre-filtering
\cite{Schaeffer1989-mu,Papadimitriou_undated-bo}).
%\todo{This paper
%https://science.sciencemag.org/content/368/6495/eaam9744.full
%shows a thousands-fold increase in speed when parallelisation
%is implemented with things like cache, architecture, etc.
%mind VS naive parallelisation as in Python, R, etc.}

Other factors that are less quantifiable that influence whether a
software tool is selected include: whether the documentation is good,
user-friendly, word-of-mouth and ``used in a similar analysis''
\cite{Loman2015-bw}. However, citation metrics may be a
useful proxy for the above.
%This may also explain the reason why
%some software continues to be used, in spite of poor relative
%performance \cite{Wadi2016-dj}.
%We note that none of these factors
%are directly linked to the top ranked features of successful software
%projects, as based upon literature surveys \cite{nasir2011critical}.

With the wide adoption of GitHub ($\approx 47\%$ of the
{\color{black}\numTools} tools included in this study could be
linked to a GitHub repository), and consequently quantifiable data on software
development such as the number of contributors to code, number of code
changes and versions is also available
\cite{ray2014large,Dozmorov:2018,mangul2018comprehensive}.

@@

\section*{Results}
We have collected relative accuracy and speed ranks for
{\color{black}\numTools~} distinct software tools. This software has
been developed for solving broad cross-section computational biology
tasks.
Each software tool was benchmarked in at least one of
{\color{black}\numBenchmarkPubs~} publications that satisfy the Boulesteix
criteria \cite{Boulesteix2013-vb}. In brief, the Boulesteix criteria
are: 1. the main focus of the article is a benchmark. 2. the authors
are reasonably neutral. 3. the test data and evaluation criteria are
sensible.

\begin{figure*}[htb!]
\includegraphics[width=\textwidth]{figure1.pdf}
\caption{\textbf{A.} A heatmap indicating the relationships between
  proposed predictors of software quality. Spearman’s rho is used to
  infer correlations between metrics such as the H- and M-indices of
  corresponding authors, number of citations, journal impact factors
  and H5 indices, the year and relative age of software and the mean
  relative rankings of software speed and accuracy. Red colours
  indicate a positive correlation, blue colours indicate a negative
  correlation. Correlations with an uncorrected P-value less than 0.05 are
  indicated with a `$\vartimes$'. The dashed rectangular area is illustrated in
  more detail in \textbf{B}, the relationship between speed and
  accuracy is shown in more detail in \textbf{Figure 2}.
  %The
  %dendrogram was computed using the default ‘heatmap.2 function in R,
  %which computes a Euclidean distance matrix and a complete-linkage
  %hierarchical clustering.
  \textbf{B.} A barplot illustrating the correlation, as measured by
  Spearman’s rho, between normalised accuracy ranks and software tool
  features that may be predictive of accuracy. In order to give an
  appreciation of the difference between the observed effect-sizes and
  significant effect sizes, we generated 1000 permuted accuracy ranks
  for each benchmark and recorded Spearman’s rho for the significant
  correlations with an uncorrected P-value less than 0.05.
  These values are marked with blues `$\vartimes$'s in the
  barplot. Significant correlations are marked with red asterisks.}
\label{fig:allfactors}
\end{figure*}

@@

We have computed the Spearman’s correlation coefficient for each
pairwise combination of the mean normalised accuracy and speed ranks,
with the year published, mean relative age (compared to software in the
same benchmarks), journal H5 metrics, the total number of citations,
the relative number of citations (compared to software in the same
benchmarks) and the maximum H- and corresponding M-indices for the corresponding
authors, version number, and numbers of commits and contributors.
The results are presented in Figure~\ref{fig:allfactors}A. We find
significant associations between most of the citation-based metrics
(journal H5, citations, relative citations, H-index and
M-index). There is also a negative correlation between the year of
publication, the relative age and many of the citation-based metrics.

Data on the
number of updates to software tools from GitHub such as the number of versions,
commits and contributors was significantly correlated with 
accuracy (respective Spearman’s rhos = {\color{black}0.13, 0.22, 0.21,} respective uncorrected P-values = {\color{black}0.0027, 0.00079, 0.0013}). 
These features were not correlated with speed however (see Figure~\ref{fig:allfactors}A).
We also found that author reputation metrics, journal impacts and the age
of tools were generally \textbf{not} correlated with either tool
accuracy or speed (see Figure~\ref{fig:allfactors}). The strongest
association was between accuracy and the relative
number of citations (Spearman’s rho = {\color{black}0.092,
P-value = 0.047}). But the effect size is only just over the
significance threshold and may be the result of multiple-testing.

@@

In order to gain a deeper understanding of the distribution of
available bioinformatic software tools on a speed versus accuracy
landscape, we ran a permutation test. The ranks extracted
from each benchmark were randomly permuted, generating 1,000
randomized speed and accuracy ranks. In the cells of a $10\times10$
grid spanning the normalised speed and accuracy ranks we computed a
Z-score for the observed number of tools in a cell, compared to the
expected distributions generated by 1,000 randomized ranks. The results
of this are shown in Figure~\ref{fig:speedaccuracy}. We identified {\color{black}18}
bins where there was a significant excess or dearth of tools. For
example, there was an excess of “slow and inaccurate” software ({\color{black}Z=2.50,
P-value=0.0063}). We find that the amount of software classed as “fast
and accurate”, “fast and inaccurate” and "slow and accurate" are at approximately the
expected proportions based upon the permutation test. 
While there are very few tools placed in the middle of the ranks on accuracy and speed.
The number of
significant results is in excess, and is not due to multiple testing, as
the probability of finding {\color{black}18} of 100 tests significant by chance is
low (P-value = {\color{black}$2.2\times 10^{-6}$}, exact binomial test).

@@

\begin{figure*}[htb!]
%\includegraphics[scale=0.55]{figure2.pdf}
\centering
\includegraphics[width=0.75\textwidth]{figure2.pdf}
\caption{A heatmap indicating the relative paucity or abundance of
  software in the range of possible accuracy and speed rankings. Red
  colours indicate an abundance of software tools in an accuracy and
  speed category, while blue colours indicate scarcity of software in
  an accuracy and speed category. The abundance is quantified using a
  Z-score computation for each bin, this is derived from 10,000 random
  permutations of the speed and accuracy ranks from each
  benchmark. Mean normalised ranks of accuracy and speed have been
  binned into 100 classes (a $10\times10$ grid) that range from
  comparatively slow/inaccurate to comparatively
  fast/accurate. Z-scores with a P-value less than 0.05 are indicated
  with a ‘$\vartimes$’.}
\label{fig:speedaccuracy}
\end{figure*}

There is a major reduction in the number of software tools that are
classed as intermediate in terms of both speed and accuracy based on
permutation tests (see Methods for details, Figure~\ref{fig:speedaccuracy}). The cells
corresponding to the four central deciles are
highlighted {\color{black}(Z = -1.9, -2.5, -2.9 and -1.1, P-values = 0.03, 0.006,
0.002 and 0.14}, respectively, reading from top to bottom, left to
right). We also tested the relative age of the software tools in
the indicated corners and central regions of the speed vs
accuracy plot. We found that the “slow and inaccurate” tools ($N=27$) were
generally published earlier than ``fast and accurate'' tools ($N=23$) {\color{black}(W = 341, P=0.005,
one-tailed Wilcoxon test)} (Figure S7).

@@

We have gathered data on the relative speeds and accuracies of
{\color{black}\numTools~} bioinformatic tools from
{\color{black}\numBenchmarkPubs} benchmarks published between
2005 and 2020. Our results provide significant support for the suggestion that there are major benefits to
the long-term support of software development \cite{siepel2019challenges}. The
finding of a strong relationship between the number of commits and code contributors to
GitHub (i.e. software updates) and accuracy, highlights the benefits of
long-term development.

Our study finds little evidence to support that impact-based metrics
have any relationship with software quality, this is unfortunate as these
are frequently cited reasons for selecting software tools
\cite{Loman2015-bw}. This implies that the high citation
rates for bioinformatic software
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a
reflection of other factors such as user-friendliness or the Matthew Effect
\cite{Lariviere2010-kx,Merton1968-cb} other than accuracy.

We found the lack of a correlation between software speed and accuracy
surprising.  The slower software tools are overrepresented at both
high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
In addition, we find an under-representation of software that has
intermediate levels of both accuracy and speed. A possible explanation
for this is that bioinformatic software tools are bound by a form of
publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
probability that a study being published is influenced by the results
it contains \cite{sterling1995publication}. The community of
developers, reviewers and editors may be unwilling to publish software
that is not highly ranked on speed or accuracy. If correct, this may
have unfortunate consequences.  We also found that slow and inaccurate
software is generally published earlier than fast and accurate tools
{\color{black}(P=0.007, one-tailed Wilcoxon test)}, therefore
comparisons were not required to publish the slow and
inaccurate tools.

@@

While we have taken pains to mitigate many issues with our analysis, nevertheless 
some limitations remain. For example, it has proven difficult to verify if the gap in medium accuracy 
and medium speed software is genuinely the result of publication bias, or due to additional factors that we have 
not taken in to account. In addition, all of the features we have used here are moving targets, e.g. as software 
tools are refined, their relative accuracies and speeds will change, the citation metrics, ages, and version control 
derived measures also change over time. We merely report a snapshot of values from 2020.  The benchmarks themselves may 
also introduce biases into the study, for example there are issues with a potential lack of independence between benchmarks 
(e.g. shared datasets, metrics and tools), there are heterogeneous measures of accuracy and speed and often unclear 
processes for including different tools.

@@

We propose that the full spectrum of software tool accuracies and
speeds serve a useful purpose to the research community. Like negative
results, if honestly reported, illustrate to the research community
that certain approaches are not practical research avenues
\cite{fanelli2012negative}. % Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
The current
practices of publishers, editors, reviewers and authors of software
tools therefore may be depriving our community of tools for building effective
and productive workflows.

@@

\textbf{Criteria for inclusion:} We are interested in using
computational biology benchmarks that satisfy Anne-Laure Boulesteix’s
(ALB) three criteria for a “neutral comparison study”
\cite{Boulesteix2013-vb}. Firstly, the main focus of the article is
the comparison and \textbf{not} the introduction of a new
tool. Secondly, the authors should be reasonably neutral, which means
that the authors should not generally have been involved in the
development of the tools included in the benchmark. Thirdly, the test
data and evaluation criteria should be sensible. This means that the
test data should be independent of data that tools have been trained
upon, and that the evaluation measures appropriately quantify correct
and incorrect predictions.

>>>>

\PaperTitle{A meta-analysis of bioinformatics software benchmarks reveals major influences on software accuracy} 
%Analysis of sixty software benchmarks shows the benefits of good software practice 
%Assessing the state of the bioinformatic software landscape with a meta-analysis of benchmarks
%Sustained software development, not citation metrics, is indicative of accurate bioinformatic software 
%

@@

%\Abstract{In the below we provide additional results for our investigation of computational biology benchmarks.}
\Abstract{Computational biology provides widely used and powerful
  software tools for testing and making inferences about biological
  data. In the face of rapidly increasing volumes of data, heuristic
  methods that trade software speed for accuracy may
  be employed.  We are have studied these trade-offs using the results of a large number of independent software benchmarks, 
  and evaluated whether external factors are indicative of accurate
  software.

We have extracted accuracy and speed ranks from independent benchmarks of
different bioinformatic software tools, and evaluated whether the
speed, author reputation, journal impact, recency and developer efforts are indicative of accuracy.

We found that software speed, author reputation, journal impact, 
number of citations and age are all unreliable predictors of software
accuracy. This is unfortunate because citations, author and journal
reputation are frequently cited reasons for selecting software
tools. However, GitHub-derived records and high version numbers show that 
the accurate bioinformatic software tools
are generally the product of many improvements over time, often from
multiple developers.

We also find
that the field of bioinformatics has a large excess of slow and inaccurate software
tools, and this is consistent across many sub-disciplines. Meanwhile, there are few tools that
are middle-of-road in terms of accuracy and speed trade-offs. We
hypothesise that a form of publication-bias influences the
publication and development of bioinformatic software. In other
words, software that is intermediate in terms of both speed and accuracy may be difficult to publish - possibly due to author,
editor and reviewer practices. This leaves an unfortunate hole in the
literature as the ideal tools may fall into this gap. For example, high accuracy tools are not always useful if years of CPU time are required, 
while high speed is not useful if the results are also inaccurate.  }

@@

\section*{Background}
Computational biology software is widely used and has produced some of
the most cited publications in the entire scientific corpus
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. These
highly-cited software tools include implementations of methods for sequence
alignment and homology inference \cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
phylogenetic analysis \cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei}, 
biomolecular structure analysis \cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj}, and
visualization and data collection \cite{Kraulis1991-lt,Berman2000-to}. 
However, the popularity of a
software tool does not necessarily mean that it is accurate or
computationally efficient, instead usability, ease of installation,
operating system support or other indirect factors may play a greater role in a software
tool's popularity. Indeed, there have been several notable incidences where
convenient, yet inaccurate software has caused considerable harm
\cite{leveson1993investigation,cummings2020regulating,Herkert:2020}.
%Fergusson covid predictions?

Progress in the biological sciences is increasingly limited by the
ability to analyse large volumes of data, therefore the
dependence of biologists on software is also increasing
\cite{Marx2013-zi}. There is an increasing reliance on technological
solutions for automating biological data generation
(e.g. next-generation sequencing, mass-spectroscopy, cell-tracking and
species tracking), therefore the biological sciences have become
increasingly dependent upon software tools for processing
large quantities of data \cite{Marx2013-zi}. As a consequence, the
computational efficiency of data processing and analysis software is
of great importance to decrease the energy, climate impact, and time costs of research
\cite{Gombiner2011-md}. Furthermore,  as datasets become
larger even small error rates can have
major impacts on the number of false inferences \cite{Storey2003-cv}.

The gold-standard for determining accuracy is for researchers independent
of individual tool development to conduct benchmarking studies, these benchmarks can serve a useful role in
reducing the over-optimistic reporting of software accuracy
\cite{Boulesteix2010-te,Jelizarow2010-zf,Weber:2019} and the self-assessment trap
\cite{Norel2011-cq,Buchka:2021}. Benchmarking typically involves the use a number of
positive and negative control datasets, so that predictions from different software tools can be
partitioned into true or false groups, allowing a variety of metrics to be
used to evaluate performance
\cite{Egan1975-nd,Hall2012-kg,Weber:2019}.
%Some benchmarks now use live, or
%frequently updated, results to indicate the latest developments in
%software performance
%\cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}.
The aim of these benchmarks is to robustly identify tools that
make acceptable compromises in terms of balancing speed with
discriminating true and false predictions, and are therefore suited
for wide adoption by the community.

For common computational biology tasks, a proliferation of
software-based solutions often exists
\cite{Felsenstein1995-ic,Altschul2013-bv,Henry2014-ut}. While
this is a good problem to have, and points to a
diversity of options from which practical solutions can be selected,
having many possible options creates a dilemma for users. In the absence of
any recent gold-standard benchmarks, how should scientific software be
selected? In the following we presume that the ``biological
accuracy'' of predictions is the most desirable feature for a software tool. Biological
accuracy is the degree to which predictions or measurements reflect
the biological truths based on expert-derived curated datasets (see Methods for the  mathematical definition used here). 
%In some
%fields biological accuracy is difficult to ascertain, for
%example, in phylogenetics it is nearly impossible in most cases to know the
%true ancestral relationships between organisms. In situations like this,
%researchers may use simulated or high-confidence datasets to assess
%software performance.

A number of possible predictors of software quality are used by the
community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. 
Some accessible, quantifiable and frequently used proxies for identifying high quality
software include: \textbf{1. Recency:} recently published software
tools may have built upon the results of past work, or be an update to
an existing tool. Therefore these may be more accurate and
faster. \textbf{2. Wide adoption:} a software tool may be widely used
because it is fast and accurate, or because it is well-supported and
user-friendly. In fact,``large user base'', ``word-of-mouth'',
``wide-adoption'', ``personal recommendation'', and ``recommendation from a
close colleague'', were frequent responses to surveys of ``how do
scientists select software?''
\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
impact:} high profile journals are run by editors and reviewers who
carefully select and curate the best manuscripts. Therefore, high
impact journals may be more likely to select manuscripts describing
good software \cite{Garfield1955-wf}. \textbf{4. Author/Group
reputation:} the key to any project is the skills of the people
involved, including maintaining a high collective intelligence
\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
consequence, an argument could be made that well respected and
high-profile authors may write better software
\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
frequently trade accuracy for speed. For example, heuristic
software such as the popular homology search tool, BLAST, compromises
the mathematical guarantee of optimal solutions for more speed
\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
interpret this fact as implying that slower software is likely to be more
accurate. But speed may also be influenced by the programming language
\cite{Fourment2008-vl}, and the level of hardware optimisation
\cite{Farrar2007-ky,Dematte2010-ph};
However, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
rapid and sensitive pre-filtering
\cite{Schaeffer1989-mu,Papadimitriou_undated-bo,leiserson2020there}).
%\todo{This paper
%https://science.sciencemag.org/content/368/6495/eaam9744.full
%shows a thousands-fold increase in speed when parallelisation
%is implemented with things like cache, architecture, etc.
%mind VS naive parallelisation as in Python, R, etc.}

%Other factors that are less quantifiable that influence whether a
%%software tool is selected include: whether the documentation is good,
%user-friendly, word-of-mouth and ``used in a similar analysis''
%\cite{Loman2015-bw}. However, citation metrics may be a
%useful proxy for the above.
%This may also explain the reason why
%some software continues to be used, in spite of poor relative
%performance \cite{Wadi2016-dj}.
%We note that none of these factors
%are directly linked to the top ranked features of successful software
%projects, as based upon literature surveys \cite{nasir2011critical}.

With the wide adoption of GitHub ($47\%$ of the
{\color{black}\numTools} tools included in this study could be
linked to a GitHub repository), and consequently quantifiable data on software
development time and intensity indicators, such as the number of contributors to code, number of code
changes and versions is available for these
\cite{ray2014large,Dozmorov:2018,mangul2018comprehensive}.

@@

\section*{Results}
We have collected relative accuracy and speed ranks for
{\color{black}\numTools~} distinct software tools. This software has
been developed for solving a broad cross-section of computational biology
tasks.
Each software tool was benchmarked in at least one of
{\color{black}\numBenchmarkPubs~} publications that satisfy the Boulesteix
criteria \cite{Boulesteix2013-vb}. In brief, the Boulesteix criteria
are: 1. the main focus of the article is a benchmark. 2. the authors
are reasonably neutral. 3. the test data and evaluation criteria are
sensible.

\begin{figure*}[htb!]
\includegraphics[width=\textwidth]{figure1.pdf}
\caption{\textbf{A.} A heatmap indicating the relationships between
  different features of bioinformatic software tools. Spearman’s rho is used to
  infer correlations between metrics such as citations based metrics, the year and relative age of publication, 
  version number, GitHub derived activity measures, and 
  the mean
  relative speed and accuracy rankings. Red colours
  indicate a positive correlation, blue colours indicate a negative
  correlation. Correlations with an P-value less than 0.05 (corrected for multiple-testing using the Benjami-Hochberg method) are
  indicated with a `$\vartimes$' symbol. The correlations with accuracy are illustrated in
  more detail in \textbf{B}, the relationship between speed and
  accuracy is shown in more detail in \textbf{Figure 2}.
  %The
  %dendrogram was computed using the default ‘heatmap.2 function in R,
  %which computes a Euclidean distance matrix and a complete-linkage
  %hierarchical clustering.
  \textbf{B.} Violin plots of Spearman's correlations between permuted accuracy ranks and 
    different software tool features. 
    The unpermuted correlations are indicated with a red asterix.
    For each benchmark, 1,000 permuted sets of accuracy and speed ranks were generated, and the ranks were normalised to lie between 0 and 1 (see Methods for details).
  Circled asterixs are significant (empirical P-value $< 0.05$, corrected for multiple-testing using the Benjami-Hochberg method).}
\label{fig:allfactors}
\end{figure*}

@@

We have computed the Spearman’s correlation coefficient for each
pairwise combination of the mean normalised accuracy and speed ranks,
with the year published, mean relative age (compared to software in the
same benchmarks), journal H5 metrics, the total number of citations,
the relative number of citations (compared to software in the same
benchmarks) and the maximum H- and corresponding M-indices for the corresponding
authors, version number, and numbers of commits and contributors.
The results are presented in Figure~\ref{fig:allfactors}A\&B. We find
significant associations between most of the citation-based metrics
(journal H5, citations, relative citations, H-index and
M-index). There is also a negative correlation between the year of
publication, the relative age and many of the citation-based metrics.

Data on the
number of updates to software tools from GitHub such as the number of versions,
commits and contributors was significantly correlated with 
software accuracy (respective Spearman’s rhos = {\color{black}0.15, 0.22, 0.21,} and respective Benjamini \& Hochberg corrected P-values = {\color{black}0.029, 0.042, 0.063}), Figure~\ref{fig:allfactors}B. 
The significance if these features was further confirmed with a permutation test (Figure~\ref{fig:allfactors}B).
These features were not correlated with speed however (see Figure~\ref{fig:allfactors}A \& Supplementary Figure ????).
We also found that reputation metrics such as citations, author and journal H-indices, and the age
of tools were generally \textbf{not} correlated with either tool
accuracy or speed (Figure~\ref{fig:allfactors}A\&B). 
%The strongest
%association was between accuracy and the relative
%number of citations (Spearman’s rho = {\color{black}0.092,
%P-value = 0.047}). But the effect size is only just over the
%significance threshold and may be the result of multiple-testing.

@@

In order to gain a deeper understanding of the distribution of
available bioinformatic software tools on a speed versus accuracy
landscape, we ran a permutation test. The ranks extracted
from each benchmark were randomly permuted, generating 1,000
randomized speed and accuracy ranks. In the cells of a $3\times3$
grid spanning the normalised speed and accuracy ranks we computed a
Z-score for the observed number of tools in a cell, compared to the
expected distributions generated by 1,000 randomized ranks. The results
of this are shown in Figure~\ref{fig:speedaccuracy}. We identified {\color{black}4}
of 9 bins where there was a significant excess or dearth of tools. For
example, there was an excess of ``slow and inaccurate'' software ({\color{black}Z=3.39,
P-value=$3.5\times 10^{-4}$}), with more moderate excess of ``slow and accurate'' and ``fast and accurate'' software (Z=2.49 \& 1.7, P=$6.3\times 10^{-3}$ \& 0.04 respectively). We find that only the ``fast and inaccurate'' extreme class is at approximately the
expected proportions based upon the permutation test (Figure 2B).

@@

\begin{figure*}[htb!]
%\includegraphics[scale=0.55]{figure2.pdf}
\centering
\includegraphics[width=0.75\textwidth]{figure2.pdf}
\caption{\textbf{A.} A heatmap indicating the relative paucity or abundance of
  software in the range of possible accuracy and speed rankings. Redder
  colours indicate an abundance of software tools in an accuracy and
  speed category, while bluer colours indicate scarcity of software in
  an accuracy and speed category. The abundance is quantified using a
  Z-score computation for each bin, this is derived from 1,000 random
  permutations of speed and accuracy ranks from each
  benchmark. Mean normalised ranks of accuracy and speed have been
  binned into 9 classes (a $3\times 3$ grid) that range from
  comparatively slow and inaccurate to comparatively
  fast and accurate. Z-scores with a P-value less than 0.05 are indicated
  with a ‘$\vartimes$’. \textbf{B.} The z-score distributions from the permutation tests (indicated with the wheat coloured violin plots) compared to 
  the z-score for the observed values for each of the corner and middle square of the heatmap.}
\label{fig:speedaccuracy}
\end{figure*}

The largest difference between the observed and expected software ranks is the reduction in the number of software tools that are
classed as intermediate in terms of both speed and accuracy based on
permutation tests (see Methods for details, Figure~\ref{fig:speedaccuracy}). The middle cell
of Figure 2A and left-most violin plot of Figure 2B highlight this extreme, (Z=-6.38,
P-value=$9.0\times 10^{-11}$).

@@

We have gathered data on the relative speeds and accuracies of
{\color{black}\numTools~} bioinformatic tools from
{\color{black}\numBenchmarkPubs} benchmarks published between
2005 and 2020. Our results provide significant support for the suggestion that there are major benefits to
the long-term support of software development \cite{siepel2019challenges}. The
finding of a strong relationship between the number of commits and code contributors to
GitHub (i.e. software updates) and accuracy, highlights the benefits of
long-term or at least intensive development.

Our study finds little evidence to support that impact-based metrics
have any relationship with software quality, which is unfortunate, as these
are frequently cited reasons for selecting software tools
\cite{Loman2015-bw}. This implies that high citation
rates for bioinformatic software
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy} is more a
reflection of other factors such as user-friendliness or the Matthew Effect
\cite{Lariviere2010-kx,Merton1968-cb} other than accuracy.

We found the lack of a correlation between software speed and accuracy
surprising.  The slower software tools are overrepresented at both
high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
In addition, there is an large under-representation of software that has
intermediate levels of both accuracy and speed. A possible explanation
for this is that bioinformatic software tools are bound by a form of
publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
probability that a study being published is influenced by the results
it contains \cite{sterling1995publication}. The community of
developers, reviewers and editors may be unwilling to publish software
that is not highly ranked on speed or accuracy. If correct, this may
have unfortunate consequences as these tools may never-the-less have further uses.
%We also found that slow and inaccurate
%software is generally published earlier than fast and accurate tools
%{\color{black}(P=0.007, one-tailed Wilcoxon test)}, therefore
%comparisons with existing software were not required to publish the slow and
%inaccurate tools.

@@

While we have taken pains to mitigate many issues with our analysis, nevertheless 
some limitations remain. For example, it has proven difficult to verify if the gap in medium accuracy 
and medium speed software is genuinely the result of publication bias, or due to additional factors that we have 
not taken in to account. In addition, all of the features we have used here are moving targets. For example, as software 
tools are refined, their relative accuracies and speeds will change, the citation metrics, ages, and version control 
derived measures also change over time. Here we report a snapshot of values from 2020.  The benchmarks themselves may 
also introduce biases into the study. For example, there are issues with a potential lack of independence between benchmarks 
(e.g., shared datasets, metrics and tools), there are heterogeneous measures of accuracy and speed and often unclear 
processes for including different tools.

@@

We propose that the full spectrum of software tool accuracies and
speeds serves a useful purpose to the research community. Like negative
results, if honestly reported this information, illustrates to the research community
that certain approaches are not practical research avenues
\cite{fanelli2012negative}. % Ioannidis2005-xh,Workman1999-au,Rivas2000-fb}.
The current
practices of publishers, editors, reviewers and authors of software
tools therefore may be depriving our community of tools for building effective
and productive workflows.

@@

\textbf{Criteria for inclusion:} We are interested in using
computational biology benchmarks that satisfy Boulesteix’s
(ALB) three criteria for a “neutral comparison study”
\cite{Boulesteix2013-vb}. Firstly, the main focus of the article is
the comparison and \textbf{not} the introduction of a new
tool. Secondly, the authors should be reasonably neutral, which means
that the authors should not generally have been involved in the
development of the tools included in the benchmark. Thirdly, the test
data and evaluation criteria should be sensible. This means that the
test data should be independent of data that tools have been trained
upon, and that the evaluation measures appropriately quantify correct
and incorrect predictions.
